---
title: "From theories to models to predictions with Bayesian model comparison."
author: "Julia Haaf"
date: May 14th, 2019
output:
  beamer_presentation:
    theme: "metropolis"
    fonttheme: "structurebold"
    fig_caption: false
    incremental: true
    slide_level: 2
    fig_crop: false
    
bibliography: lab.bib
csl: "apa6.csl"
    
header-includes   :
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{setspace}
   - \usepackage{pcl} 
   - \usepackage{color,xcolor}
   - \definecolor{green}{rgb}{0,.30,0}
   - \definecolor{red}{rgb}{.7,0,0}
   - \definecolor{blue}{rgb}{0,0,.8}
   - \usepackage{wasysym}
   - \usepackage{cancel}
   - \usepackage{multirow}
   - \titlegraphic{\centering \includegraphics[height=1.5cm]{pics/UvAlogo.png} \vspace{2cm}}

   - \newcommand\Fontvi{\fontsize{6}{7.2}\selectfont}
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')

library('msm')
library('knitr')
library("MASS")
library("boot")
library("plotrix")
library("reshape2")
library("shape")
library("MCMCpack")
library("curl")
library("plyr")
library("psych")
library("diagram")
library("RColorBrewer")
library("spatialfil")
library("ggplot2")
library("papaja")
library("mvtnorm")
library("tmvtnorm")
library("cowplot")
library("gridExtra")
library("BayesFactor")
library("rvest")
library("Hmisc")
library("devtools")
library("readxl")
library('tidyr')
library("foreign")
library("dplyr")
library("car")

SourceURL <- "https://raw.githubusercontent.com/Joe-Hilgard/Anderson-meta/master/table_managing_functions.R"
source_url(SourceURL)

SourceURL <- "https://raw.githubusercontent.com/PerceptionAndCognitionLab/ctx-indiff/public/shared/functions/getdata.R"
source_url(SourceURL)

polyCI <- function(upper, lower, col){
  len <- length(upper)
  polygon(x = c(rev(1 : len), 1 : len)
          , y = c(rev(lower), upper)
          , col = col
          , border = NA)
}

plot_freq_bayes <- function(ybar
                            , exp_var = "exp"
                            , cond_var = "cond"
                            , congruent = 0
                            , thet.n = theta.n
                            , main = ""
){
  
  #####################FREQUENTIST PLOT#############################################
  exp_num <- unique(exp_var)
  delta <- data.frame(exp = exp_num, delta = NA)
  delta$delta <- ybar[, 2] - ybar[, 1]
  delta$thn <- colMeans(thet.n) * 1000
  delta[, ncol(delta)+(1:2)] <- t(apply(thet.n, 2, function(x) quantile(x, c(.025, .975))))
  delta_i <- delta[order(delta$delta),]
  delta_i$delta <- delta_i$delta *1000
  
  plot(x = 1: nrow(delta_i)
       , y = delta_i$delta
       , ylim = c(min(delta_i$delta), max(delta_i$delta))
       , type = "l"
       , col = "gray50"
       , ylab = expression(Effect ~ parameter ~ theta[i])
       , xlab = "Studies"
       , frame.plot = FALSE
       , xaxt = 'n'
       , yaxt = 'n'
  )
  title(main, line = -1)
  axis(side = 1
       , at = c(1, nrow(delta_i))
  )
  axis(side = 2
  )
  polyCI(upper = delta_i[,ncol(delta)] * 1000, lower = delta_i[,ncol(delta)-1] * 1000, col = "gray85")
  
  abline(h = 0, col = "gray40")
  
  lines(delta_i$delta
         , col = "gray65"
        , lwd = 2)
  
  points(delta_i$thn
         , col = "darkslateblue"
         , pch = 19
         )
}
```

# How can we test psychological theories?

## Theoretical Statements

- Levels of Processing: *Processing words semantically increases memory performance as compared to more shallow processing.*
- Stroop: *Inhibition leads to a slow-down for incongruent items compared to congruent items.*
- Positive psychology: *Positive psychology interventions improve life satifaction.*

\vspace*{.5cm}

\centering \includegraphics[width = 100px]{pics/theory.png}

## Theoretical Statements

- Each statement describes constraint in the world.
- None of these statements may be tested with data.
- None has specification of sampling noise or measurement errors.
- And that is appropriate.

\vspace*{.5cm}

\centering \includegraphics[width = 100px]{pics/theory.png}

## Statistical Models

- Models instantiate the critical constraints from theory.
- Models specify sample noise.
- Models can predict data.

```{r, eval = F}
M  <- matrix(nrow = 6, ncol = 6, byrow = TRUE, data = 0)
M[2, 1] <- ""
M[3,1]<-""
M[4,1]<- ""
M[5,2]<-""
M[5,3]<-""    
M[5, 4] <- "Analysis"
M[6,5] <-"Interpretation"

ArrM <- matrix(nrow = 6, ncol = 6, byrow = TRUE, data = 0)

ArrM[2, 1] <- .5
ArrM[3,1]<-.33
ArrM[4,1]<- .25
ArrM[5,2]<-.5
ArrM[5,3]<- .5    
ArrM[5, 4] <- .5
ArrM[6,5] <-.5

col=c("lightskyblue1","thistle1","thistle1","thistle1","darkseagreen1","lightskyblue1")

pos <- cbind (c(0.17, 0.17, 0.27, 0.43, 0.79, 0.79), c(0.89, 0.5, 0.3, 0.1, 0.4, 0.89))
pp <- plotmat(M, pos = pos, name = c("Theoretical\n Constraint", "Model:\n Constraint I", "Model:\n Constraint II", "Model:\n Constraint III","Statistical\n Inference", "Inference\n About Theory"),box.col=col,
              lwd = 1, box.lwd = 2, cex.txt = 1.3, 
              box.size = 0.16, box.type = "rect", box.prop = 0.25, curve = 0
              , arr.type = "triangle"
              , arr.lwd = 1.3
              , arr.pos = ArrM
              , shadow.size = 0
              , box.cex = 1.4)
```

\vspace*{1cm}

\begin{figure}
\hfill
\includegraphics[width = 80px]{pics/theory.png}
\end{figure}



## How do we represent numbers internally?
\framesubtitle{How do we represent numbers internally?}

\vspace*{.3cm}

\centering \includegraphics[width = 150px]{pics/numbers.png}

@Rouder:etal:2005a

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

1. Analog representation.

\vspace*{.7cm}

\centering \includegraphics[width = 250px]{pics/skala.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

>1. Analog representation.
>2. Propositional representation.

\vspace*{.4cm}

\centering \includegraphics[width = 290px]{pics/verhaeltnis.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

>1. Analog representation.
>2. Propositional representation.
>3. Priming + spreading activation.

\vspace*{.2cm}

\centering \includegraphics[width = 250px]{pics/optinussprime.png}

<!-- ## How do we represent numbers internally? -->
<!-- \framesubtitle{Theoretical Positions:} -->

<!-- \vspace*{.3cm} -->

<!-- >1. Everyones uses analog representation. -->
<!-- >2. Everyone uses propositional representation. -->
<!-- >3. Everyone uses priming + spreading activation. -->
<!-- >4. None of the above. -->

<!-- \vspace*{.2cm} -->

<!-- \centering \includegraphics[width = 130px]{pics/whoot.png} -->

## Default statistical model

## Default statistical model

```{r, cache = T}
source('../../papers/submission/R-scripts/ld5.R')

dat=clean()
dat$side <- factor(ifelse(dat$stim >= 3, "g", "l"))
dat$proxi <- factor(ifelse(dat$stim %in% c(0, 5), 3, ifelse(dat$stim %in% c(2, 3), 1, 2)))
res <- anovaBF(rt ~ side*proxi, dat)
```

- ANOVA: 2 (direction: greater vs. less) x 3 (proximity: 2 or 8 vs. 3 or 7 vs. 4 or 6)
- Theoretically we are concerned with the second main effect.
- Main effect of proximity: Bayes factor of $10^{40}$-to-1 in favor of the effect.
- Bayes factor of at least $`r printnum(extractBF(res)[2, 1]/extractBF(res)[3, 1])`$-to-1 against any other model.

\vspace*{.5cm}

\begin{figure}
\hfill
\includegraphics[width = 75px]{pics/numbers.png}
\end{figure}

## Default statistical model

But was does it mean to have significant main effect?

```{r, fig.asp = .4, fig.align='center'}
layout(matrix(1:2, ncol = 2))

predm1 <- c(550, 610, 650, 640, 600, 540)

plot(0, 0, ylim = c(500, 700), xlim = c(.5, 6.5), axes = F, ylab="Response Time (ms)",xlab='Digit')
lines(1:3,predm1[1:3], lwd = 2)
points(1:3,predm1[1:3],pch=21,bg='firebrick1',cex=1.2)
lines(4:6,predm1[4:6], lwd = 2)
points(4:6,predm1[4:6],pch=21,bg='firebrick1',cex=1.2)
axis(2, at = seq(550, 650, 50))
axis(1,at=1:3,lab=c(2:4))
axis(1,at=4:6,lab=c(6:8))

predm1 <- c(650, 610, 550, 540, 600, 640)

plot(0, 0, ylim = c(500, 700), xlim = c(.5, 6.5), axes = F, ylab="Response Time (ms)",xlab='Digit')
lines(1:3,predm1[1:3], lwd = 2)
points(1:3,predm1[1:3],pch=21,bg='firebrick1',cex=1.2)
lines(4:6,predm1[4:6], lwd = 2)
points(4:6,predm1[4:6],pch=21,bg='firebrick1',cex=1.2)
axis(2, at = seq(550, 650, 50))
axis(1,at=1:3,lab=c(2:4))
axis(1,at=4:6,lab=c(6:8))
```


## Number representation (Rouder et al., 2005)

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r plot-ld5, fig.width=6, fig.height=4.5, fig.align='center'}
par(mar = c(4, 6, 0, 2), mgp = c(2, .7, 0))
plotMean(dat,'')
```

\column{.1\textwidth}
\end{columns}


## Problem for preregistration

- Typical procedure: Test + plot.
- Test main effects and interactions (preregister).
- Plot and check whether the effect is in the right direction.
- Allows for unnecessary flexibility and posthoc reasoning.
- Specifying more constraint models may lead to more evidence for the theoretical predictions.

## Ordinal models across people

\vspace*{1cm}

\centering \includegraphics[width = 175px]{pics/arrows.png}

## Does everyone truly represent numbers as analog quantities?

- *Everyone has the same number representation* $\rightarrow$ common mechanism.
- *Mixed number representation* $\rightarrow$ more complex. What determines who has what type of number representation?

## Theoretical positions as ordinal models

\vspace*{.2cm}

```{r, fig.width=6, fig.asp=.85, fig.align='center'}
layout(matrix(ncol=2, 1:4, byrow=T))
par(cex=1.1,mgp=c(2,.7,0)
    , mar = c(.5, 2, 2, .5))
myCol=c('lightblue','lavender','lightgreen')

names=c(4:2,6:8)
o=c(1,4,2,5,3,6)
M=matrix(nrow=length(names),byrow=F,ncol=length(names),data=0)
M[1,2]=M[2,3]=''
M[4,5]=M[5,6]=''
plotmat(t(M[o,o]),pos=c(2,2,2),name=names[o],curve=0,box.type="round",box.size=.07,box.prop=1,box.col=rep(myCol,each=2),arr.length=0)
axis(2, at=c(.2,.8), lab=c("Faster","Slower"),cex=1)
# box()
mtext(side=3,adj=.5,line=-.5,"Analog-Representation",cex=1.1)

names=c(2:4,6:8)
M=matrix(nrow=length(names),byrow=F,ncol=length(names),data=0)
M[1,2]=M[2,3]=''
M[4,5]=M[5,6]=''
pos=matrix(nrow=6,ncol=2)
pos[,1]=c(.1,.4,.7,.3,.6,.9)
pos[,2]=c(.65,.65,.65,.35,.35,.35)
plotmat(M,pos,name=names,curve=0,box.type="round",box.size=.04,box.prop=1,box.col=c(rev(myCol),myCol),arr.length=0,arr.type='triangle')
# box()
mtext(side=3,adj=.5,line=-.5,"Propositional Representation",cex=1.1)


names=c(2:4,8:6)
o=c(1,4,2,5,3,6)
M=matrix(nrow=length(names),byrow=F,ncol=length(names),data=0)
M[1,2]=M[2,3]=''
M[4,5]=M[5,6]=''
plotmat(t(M[o,o]),pos=c(2,2,2),name=names[o],curve=0,box.type="round",box.size=.06,box.prop=1,box.col=myCol[c(3,3,2,2,1,1)],arr.length=0)
# box()
mtext(side=3,adj=.5,line=-.5,"Priming + Spreading Activation",cex=1.1)
axis(2, at=c(.2,.8), lab=c("Faster","Slower"),cex=1)

names=c(2:4,6:8)
M=matrix(nrow=length(names),byrow=F,ncol=length(names),data=0)
# M[1,2]=M[2,3]=''
# M[4,5]=M[5,6]=''
pos=matrix(nrow=6,ncol=2)
pos[,1]=c(.1,.4,.7,.3,.6,.9)
pos[,2]=c(.65,.65,.65,.35,.35,.35)
plotmat(M,pos,name=names,curve=0,box.type="round",box.size=.04,box.prop=1,box.col=c(rev(myCol),myCol),arr.length=0,arr.type='triangle')
# box()
mtext(side=3,adj=.5,line=-.5,"None of the above",cex=1.1)
```


## "Does Everyone" analysis 

\vspace*{.5cm}

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r plot-ld5-all, fig.width=6, fig.height=4.5, fig.align='center'}
means <- with(dat, tapply(rt, list(sub, stim), mean))
matplot(t(means), pch = 20, col = "gray50", axes = F
        , xlab = "Digit", ylab = "Response Time (ms)")
axis(1,at=1:3,lab=c(2:4))
axis(1,at=4:6,lab=c(6:8))
axis(2, at = seq(500, 1000, 250))
for(i in 1:nrow(means)){
  lines(1:3, means[i, 1:3], col = adjustcolor("gray50", .5), lwd = 1.2)
  lines(4:6, means[i, 4:6], col = adjustcolor("gray50", .5), lwd = 1.2)
}
lines(1:3, colMeans(means[,1:3]), lwd = 2)
points(1:3, colMeans(means[,1:3]), pch=21, bg='cornflowerblue', cex=1.2)
lines(4:6, colMeans(means[,4:6]), lwd = 2)
points(4:6, colMeans(means[,4:6]), pch=21, bg='cornflowerblue', cex=1.2)
```

\column{.1\textwidth}
\end{columns}

## "Does Everyone" analysis

\vspace*{1cm}

```{r, fig.width=8, fig.height=5}
myMeans=tapply(dat$rt,list(dat$sub,dat$stim),mean)
diffs=apply(myMeans,1,diff)
a1=diffs[1,]
b1=diffs[2,]
a2=-diffs[4,]
b2=-diffs[5,]
combo=cbind(a1,b1,a2,b2)
plotIndv(combo,'',ylim=c(-60,200)
         , xlabel = c(expression(paste(d[1]))
             ,expression(paste(d[2]))
             ,expression(paste(d[3]))
             ,expression(paste(d[4]))))
lines(colMeans(combo), col = "orangered2", lwd = 1.5)
points(colMeans(combo), col = "orangered2", pch = 19)
```

## Results for "Does Everyone" analysis

\vspace*{1cm}

```{r, fig.height=4.5, fig.width=8}
layout(matrix(1:2, ncol = 2))
myMeans=tapply(dat$rt,list(dat$sub,dat$stim),mean)
diffs=apply(myMeans,1,diff)
a1=diffs[1,]
b1=diffs[2,]
a2=-diffs[4,]
b2=-diffs[5,]
combo=cbind(a1,b1,a2,b2)
plotIndv(combo,'Observed Effects',ylim=c(-60,200)
         , xlabel = c(expression(paste(d[1]))
             ,expression(paste(d[2]))
             ,expression(paste(d[3]))
             ,expression(paste(d[4]))))
lines(colMeans(combo), col = "orangered2", lwd = 1.2)
points(colMeans(combo), col = "orangered2", pch = 20)

load('../../papers/submission/ld5.Rdata')
plotIndv(out$samples,'Estimated Effects',ylim=c(-60,200))
lines(colMeans(out$samples), col = "orangered2", lwd = 1.2)
points(colMeans(out$samples), col = "orangered2", pch = 20)
```

## Results for "Does Everyone" analysis

\vspace*{.2cm}

- Preferred model: **Analog representation** model

- Preferred `r printnum(out$BF[1])`-to-1 over the **None of the above** model

- Preferred `r "$3 \\times 10^{55}$"`-to-1 over the **Propositional representation** model

- Bayes factor for **Priming + spreading activation** model cannot be estimated 

\centering \includegraphics[width = 150px]{pics/skala.png}

##Take-away for theory development

1. Psychological theories typically provide ordinal constraint.
2. Developing ordinal models helps us determine predicted patterns *before* collecting data.
3. Bayesian model comparison gives an accessible framework to test ordinal models.
4. Theoretical ordinal constraints may account for everyone.
5. "Everyone Does" models are parsimonious descriptions of common mechanisms.
6. To assess common mechanisms sample noise has to be taken into account.

\vspace*{.1cm}

\begin{figure}
\centering
\includegraphics[width = 70px]{pics/theory.png}
\end{figure}

## Thank You!

Further Recources for the Pragmatic Psychologist

>- Accessible book to get started with ordinal models: @Hoijtink:2012
>- Software packages: `bain` or `BayesFactor` for `R` users
>- "Does everyone" analysis: @Haaf:Rouder:2017 and @Haaf:etal:2018
>- Presentation and all code is available at: \color{blue} \href{https://tinyurl.com/OrdinalConstraint}{tinyurl.com/OrdinalConstraint}

\begin{figure}
\hfill
\includegraphics[width = 60px]{pics/allein.png}
\end{figure}


## References

\scriptsize 
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}



---
nocite: |
  @Hoijtink:2012,
  @Rouder:etal:2012, 
  @Rouder:etal:2005a
...

