---
title: "Theories as Ordinal Constraint"
author: "Julia M. Haaf"
csl: apa6.csl
output:
  beamer_presentation:
    slide_level: 2
    keep_tex: true
    theme: "Boadilla"
    colortheme: "dove"
    fonttheme: "structuresmallcapsserif"
  ioslides_presentation:
    css: styles.css
    logo: MUlogoRGB.png
    transition: faster
classoption: t
header-includes:
- \usepackage{bm}
- \usepackage{pcl}
- \usepackage{amsmath}
- \usepackage{setspace}
- \usepackage{multirow}
- \usepackage{xcolor}
- \usepackage{adjustbox}
- \titlegraphic{\centering \includegraphics[height=2cm, width=5.5cm]{pics/MULogoBig.png}}
- \setbeamerfont{author}{series=\bfseries,parent=structure}
- \setbeamerfont{frametitle}{size=\Large}
- \setbeamerfont{title}{size = \Huge}

subtitle: A Bayes Factor Approach
bibliography: lab.bib
---

```{r include = F}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(dev = 'pdf')

library('msm')
library('knitr')
library("MASS")
library("boot")
library("scam")
library("plotrix")
library("reshape2")
library("shape")
library("MCMCpack")
library("curl")
library("plyr")
library("dplyr")
library("psych")
library("diagram")
library("RColorBrewer")
library("spatialfil")
library("ggplot2")
library("papaja")
library("mvtnorm")
library("tmvtnorm")
library("cowplot")
library("gridExtra")
library("BayesFactor")
library("rvest")
library("Hmisc")
library("devtools")
library("readxl")


SourceURL <- "https://raw.githubusercontent.com/Joe-Hilgard/Anderson-meta/master/table_managing_functions.R"
source_url(SourceURL)

SourceURL <- "https://raw.githubusercontent.com/PerceptionAndCognitionLab/ctx-indiff/public/shared/functions/getdata.R"
source_url(SourceURL)

polyCI <- function(upper, lower, col){
  len <- length(upper)
  polygon(x = c(rev(1 : len), 1 : len)
          , y = c(rev(lower), upper)
          , col = col
          , border = NA)
}

plot_freq_bayes <- function(ybar
                            , exp_var = "exp"
                            , cond_var = "cond"
                            , congruent = 0
                            , thet.n = theta.n
                            , main = ""
){
  
  #####################FREQUENTIST PLOT#############################################
  exp_num <- unique(exp_var)
  delta <- data.frame(exp = exp_num, delta = NA)
  delta$delta <- ybar[, 2] - ybar[, 1]
  delta$thn <- colMeans(thet.n) * 1000
  delta[, ncol(delta)+(1:2)] <- t(apply(thet.n, 2, function(x) quantile(x, c(.025, .975))))
  delta_i <- delta[order(delta$delta),]
  delta_i$delta <- delta_i$delta *1000
  
  plot(x = 1: nrow(delta_i)
       , y = delta_i$delta
       , ylim = c(min(delta_i$delta), max(delta_i$delta))
       , type = "l"
       , col = "gray50"
       , ylab = expression(Effect ~ parameter ~ theta[i])
       , xlab = "Studies"
       , frame.plot = FALSE
       , xaxt = 'n'
       , yaxt = 'n'
  )
  title(main, line = -1)
  axis(side = 1
       , at = c(1, nrow(delta_i))
  )
  axis(side = 2
  )
  polyCI(upper = delta_i[,ncol(delta)] * 1000, lower = delta_i[,ncol(delta)-1] * 1000, col = "gray85")
  
  abline(h = 0, col = "gray40")
  
  lines(delta_i$delta
         , col = "gray65"
        , lwd = 2)
  
  points(delta_i$thn
         , col = "darkslateblue"
         , pch = 19
         )
}
```

## Testing Theories Using Ordinal Constraint

\vspace*{1cm}

>- Is there a *true* effect?
>- Example: Inhibition ($\rightarrow \rightarrow \boldsymbol{\leftarrow} \rightarrow \rightarrow$ vs. $\leftarrow \leftarrow \boldsymbol{\leftarrow} \leftarrow \leftarrow$)
>- Is the *true* effect in the expected direction?
>- **Idea: Stating theory as multiple, simultaneous ordinal constraint**

\vspace*{.5cm}

\centering \includegraphics[width = 50px]{pics/Direction.png}

## A psychologist's favorite design
\framesubtitle{With Ugly Bugs}

\vspace*{.5cm}

\centering \includegraphics[width = 120px]{pics/newBug.png}

## A psychologist's favorite design
\framesubtitle{Ryan, Wilde, \& Crist (2013)}

\vspace*{1cm}

\centering
\begin{tabular}{llcc}
\hline
& & \multicolumn{2}{c}{\bf{Disgust}} \\ \cline{2-4}
& & Low & High\\ \hline
\multirow{2}{*}{\bf{Fear}} & Low & Low/Low & Low/High \\
& High & High/Low & High/High    \\ \hline
\end{tabular}

\vspace*{.8cm}

\bf{"How willing are you to kill this bug?"}

\vspace*{.4cm}

\centering \includegraphics[origin=c, angle = 90, width = 250px]{pics/newBug.png}


## Systems of Orders with Bugs

\vspace*{.5cm}

```{r fig-1, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Orders with Bugs

\vspace*{.5cm}

```{r, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

# #Compensatory
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# # plotmat(M
# #         , pos = c(3, 1)
# #         , name = names[c(2, 4, 3, 1)]
# #         , curve = 0
# #         , box.type="round"
# #         , box.size=.06
# #         , box.prop=.8
# #         , box.col = myCol
# #         , arr.length=0
# #         , box.cex = 1.2
# #         , relsize = 1
# #         )
# plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Compensatory"), side = 3, cex = 1.5)

#Compensatory strict

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)

mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(2, 2)
#         , name = names[c(3, 4, 1, 2)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(2, 2)
#         , name = names[c(2, 4, 1, 3)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

# plotmat(M
#         , pos = c(1, 2, 1)
#         , name = names[c(4, 2, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Orders with Bugs

\vspace*{.5cm}

```{r, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

# #Compensatory
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol[c(2, 4, 3, 1)]
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
# # plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Compensatory"), side = 3, cex = 1.5)

#Compensatory strict

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(3, 1)
        , name = names[c(2, 4, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(2, 2)
#         , name = names[c(3, 4, 1, 2)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(2, 2)
#         , name = names[c(2, 4, 1, 3)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

# plotmat(M
#         , pos = c(1, 2, 1)
#         , name = names[c(4, 2, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Orders with Bugs

\vspace*{.5cm}

```{r, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

# #Compensatory
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol[c(2, 4, 3, 1)]
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
# # plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Compensatory"), side = 3, cex = 1.5)

#Compensatory strict

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(3, 1)
        , name = names[c(2, 4, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(2, 2)
        , name = names[c(3, 4, 1, 2)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(3, 4, 1, 2)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

# plotmat(M
#         , pos = c(2, 2)
#         , name = names[c(2, 4, 1, 3)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

# plotmat(M
#         , pos = c(1, 2, 1)
#         , name = names[c(4, 2, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Orders with Bugs

\vspace*{.5cm}

```{r, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

# #Compensatory
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol[c(2, 4, 3, 1)]
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
# # plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Compensatory"), side = 3, cex = 1.5)

#Compensatory strict

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(3, 1)
        , name = names[c(2, 4, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(2, 2)
        , name = names[c(3, 4, 1, 2)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(3, 4, 1, 2)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(2, 2)
        , name = names[c(2, 4, 1, 3)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 1, 3)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

# plotmat(M
#         , pos = c(1, 2, 1)
#         , name = names[c(4, 2, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Orders with Bugs

\vspace*{.5cm}

```{r, fig.height=5.5, fig.width=9}
par(mar = c(0, 0, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
names <- c("LL", "LH", "HL", "HH")
o <- 1:4

#Null model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(4)
        , name = names
        , curve = 0
        , box.type="round"
        , box.size=.02
        , box.prop=1
        , box.col = myCol
        , arr.length=0
        , box.cex = 1.2
        , shadow.size = 0.007
        )
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.5, line = -2)

#Additive model

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

# #Compensatory
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol[c(2, 4, 3, 1)]
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         )
# # plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Compensatory"), side = 3, cex = 1.5)

#Compensatory strict

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[2, 3] <- ''
M[1, 4] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(3, 1)
        , name = names[c(2, 4, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

#Fear only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(2, 2)
        , name = names[c(3, 4, 1, 2)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(3, 4, 1, 2)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.5)

#Disgust only

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(2, 2)
        , name = names[c(2, 4, 1, 3)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(2, 4, 1, 3)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.5)

#Unconstrained

M <- matrix(nrow = length(o), ncol = length(o), data=0)

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.5)
```

## Systems of Order Analysis with Bugs
\framesubtitle{Data}

\vspace*{.5cm}

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r get-data-bugs, child="anovaAnalysis.Rmd", cache = T}
```

```{r plot-bugs, fig.width=5, fig.height=3.5, fig.align='center'}
# apa_lineplot(datl, id = "Subject", factors = c("disgust", "fear"), dv = "value")
# with(datl, tapply(value, list(fear, disgust), mean, na.rm = T))

apa_lineplot(datl, id = "Subject", factors = c("disgust", "fear"), dv = "value"
            , ylab = "Willingness to Kill"
            , xlab = "Disgust"
            , args_x_axis = list(labels = c("Low", "High"))
            , ylim = c(4, 9)
            , args_legend = list(legend = c("Low", "High")
                                 , title = "Fear", x = 1.6, y = 6.5, lwd = 1.5)
            , args_points = list(col = c("orangered2", "royalblue4")
                                 , bg = c("orangered2", "royalblue4"))
            , args_lines = list(col = c("orangered2", "royalblue4"), lwd = 1.5)
            , args_error_bars = list(col = c("orangered2", "royalblue4"))
            , jit = 0)
```

\column{.1\textwidth}
\end{columns}

## Systems of Order Analysis with Bugs
\framesubtitle{Results}

\vspace*{.5cm}

- Preferred model: **Consistent postitive** model ($M_1$)

- Preferred `r printnum(bf1u/bf3u)`-to-1 over the **positive equality** model ($M_2$)
- Preferred `r printnum(bf1u/bf4u)`-to-1 over the **fear only** model ($M_3$)
- Preferred `r printnum(bf1u)`-to-1 over the **unconstrained model** ($M_5$)
- Preferred `r paste(round(bf1u/bf5u))`-to-1 over the **disgust only** model ($M_4$)
- Preferred `r paste(round(bf1u/bf0u))`-to-1 over the **null** model ($M_0$)

```{r bf-bugs, fig.height=4.5, eval = F}
par(mar = c(2, .1, 2.5, 0))
layout(matrix(c(0, 1, 1, 1, 1, 0
                , 2, 2, 3, 3, 4, 4
                , 0, 5, 5, 6, 6, 0)
              , ncol = 6, byrow = T))

#Null model

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[0]: ~ "Null"), side = 3, cex = 1.9)
mtext(bquote(BF["0u"] ~ "=" ~ .(round(bf0u, 4))), side = 3, cex = 1.7, line = -5)

#Additive model

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(bold(M[1]: ~ "Consistent +")), side = 3, cex = 1.9)
mtext(bquote(BF["1u"] ~ "=" ~ .(round(bf1u, 2))), side = 3, cex = 1.7, line = -5)
box(col = "orangered2", lwd = 1.5)

# #Compensatory
# 
# plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "Weak Positive"), side = 3, cex = 1.5)
# mtext(bquote(BF["2u"] ~ "=" ~ .(round(bf2u, 2))), side = 3, cex = 1.2, line = -5)

#Compensatory strict

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.9)
mtext(bquote(BF["2u"] ~ "=" ~ .(round(bf3u, 2))), side = 3, cex = 1.7, line = -5)

#Fear only

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Fear Only"), side = 3, cex = 1.9)
mtext(bquote(BF["3u"] ~ "=" ~ .(round(bf4u, 2))), side = 3, cex = 1.7, line = -5)
box(col = "gray99", lwd = 1.5)

#Disgust only

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "Disgust Only"), side = 3, cex = 1.9)
mtext(bquote(BF["4u"] ~ "=" ~ .(round(bf5u, 4))), side = 3, cex = 1.7, line = -5)

#Unconstrained

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[5]: ~ "Unconstrained"), side = 3, cex = 1.9)
mtext(bquote(BF["uu"] ~ "=" ~ 1), side = 3, cex = 1.7, line = -5)
```

## Preferred model

\vspace*{1cm}

\begin{columns}
\column{.3\textwidth}
\column{.4\textwidth}

```{r, fig.width=6, fig.height=5, fig.align='center'}
M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

plotmat(M
        , pos = c(1, 2, 1)
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)
# 
# M <- matrix(nrow = length(o), ncol = length(o), data=0)
# M[1, 2] <- ''
# M[2, 3] <- ''
# M[1, 4] <- ''
# M[2, 4] <- ''
# M[3, 4] <- ''
# 
# plotmat(M
#         , pos = c(3, 1)
#         , name = names[c(2, 4, 3, 1)]
#         , curve = 0
#         , box.type="round"
#         , box.size=.06
#         , box.prop=.8
#         , box.col = myCol[c(2, 4, 3, 1)]
#         , arr.length=0
#         , box.cex = 1.2
#         , relsize = 1
#         , shadow.size = 0.007
#         )
# # plot(0, type='n', axes=FALSE, ann=FALSE)
# mtext(expression(M[2]: ~ "+ Equality"), side = 3, cex = 1.5)

```

\column{.3\textwidth}
\end{columns}

##Take-away for the Pragmatic Psychologist

\vspace*{1cm}

>- You know a lot about ordinal predictions from theory!
>- Bayesian model comparison gives an accessible framework to test *your* predictions
>- Bayesian analysis means you have to add value even after running your analysis
>- Are you convinced by a Bayes factor of 3-to-1?

\vspace*{1.5cm}

\begin{figure}
\hfill
\includegraphics[width = 80px]{pics/psychology.png}
\end{figure}

## A Step Further: Ordinal Constraint Across People

\vspace*{1cm}

\centering \includegraphics[width = 175px]{pics/arrows.png}

## A step further
\framesubtitle{Ordinal constraint across people}

\vspace*{1cm}

>- Stimulus strength: Does everyone truly detect bright lights faster than dim ones? (Most likely)

>- Handedness: Does everyone have a true right handed advantage? (Clearly no)

>- **Symbolic distance: Does everyone truly represent numbers as analog quantities? (IDK)**

## Symbolic Distance
\framesubtitle{How do we represent numbers internally?}

\vspace*{.3cm}

\centering \includegraphics[width = 150px]{pics/numbers.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

1. Everyones uses analog representation.

\vspace*{.7cm}

\centering \includegraphics[width = 250px]{pics/skala.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

1. Everyones uses analog representation.
2. Everyone uses propositional representation.

\vspace*{.4cm}

\centering \includegraphics[width = 290px]{pics/verhaeltnis.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

1. Everyones uses analog representation.
2. Everyone uses propositional representation.
3. Everyone uses priming + spreading activation.

\vspace*{.2cm}

\centering \includegraphics[width = 250px]{pics/optinussprime.png}

## How do we represent numbers internally?
\framesubtitle{Theoretical Positions:}

\vspace*{.3cm}

1. Everyones uses analog representation.
2. Everyone uses propositional representation.
3. Everyone uses priming + spreading activation.
4. None of the above.

\vspace*{.2cm}

\centering \includegraphics[width = 130px]{pics/whoot.png}

## Results for "Does Everyone" Analysis 
\framesubtitle{Rouder, Lu, Speckman, Sun, \& Jiang (2005)}

\vspace*{.5cm}

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r plot-ld5, fig.width=6, fig.height=4.5, fig.align='center'}
source('../../papers/short/R-scripts/ld5.R')

dat=clean()
plotMean(dat,'')
```

\column{.1\textwidth}
\end{columns}

## Results for "Does Everyone" Analysis 
\framesubtitle{Rouder, Lu, Speckman, Sun, \& Jiang (2005)}

\vspace*{.5cm}

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r plot-ld5-all, fig.width=6, fig.height=4.5, fig.align='center'}
means <- with(dat, tapply(rt, list(sub, stim), mean))
matplot(t(means), pch = 20, col = "gray50", axes = F
        , xlab = "Digit", ylab = "Response Time (ms)")
axis(1,at=1:3,lab=c(2:4))
axis(1,at=4:6,lab=c(6:8))
axis(2, at = seq(500, 1000, 250))
for(i in 1:nrow(means)){
  lines(1:3, means[i, 1:3], col = adjustcolor("gray50", .5), lwd = 1.2)
  lines(4:6, means[i, 4:6], col = adjustcolor("gray50", .5), lwd = 1.2)
}
lines(1:3, colMeans(means[,1:3]), lwd = 2)
points(1:3, colMeans(means[,1:3]), pch=21, bg='cornflowerblue', cex=1.2)
lines(4:6, colMeans(means[,4:6]), lwd = 2)
points(4:6, colMeans(means[,4:6]), pch=21, bg='cornflowerblue', cex=1.2)
```

\column{.1\textwidth}
\end{columns}

## Results for "Does Everyone" Analysis
\framesubtitle{Rouder et al. (2005)}

\vspace*{1cm}

```{r, fig.width=8, fig.height=5}
myMeans=tapply(dat$rt,list(dat$sub,dat$stim),mean)
diffs=apply(myMeans,1,diff)
a1=diffs[1,]
b1=diffs[2,]
a2=-diffs[4,]
b2=-diffs[5,]
combo=cbind(a1,b1,a2,b2)
plotIndv(combo,'',ylim=c(-60,200)
         , xlabel = c(expression(paste(d[1]))
             ,expression(paste(d[2]))
             ,expression(paste(d[3]))
             ,expression(paste(d[4]))))
lines(colMeans(combo), col = "orangered2", lwd = 1.5)
points(colMeans(combo), col = "orangered2", pch = 19)
```

## Results for "Does Everyone" Analysis

\vspace*{1cm}

```{r, fig.height=4.5, fig.width=8}
layout(matrix(1:2, ncol = 2))
myMeans=tapply(dat$rt,list(dat$sub,dat$stim),mean)
diffs=apply(myMeans,1,diff)
a1=diffs[1,]
b1=diffs[2,]
a2=-diffs[4,]
b2=-diffs[5,]
combo=cbind(a1,b1,a2,b2)
plotIndv(combo,'Observed Effects',ylim=c(-60,200)
         , xlabel = c(expression(paste(d[1]))
             ,expression(paste(d[2]))
             ,expression(paste(d[3]))
             ,expression(paste(d[4]))))
lines(colMeans(combo), col = "orangered2", lwd = 1.2)
points(colMeans(combo), col = "orangered2", pch = 20)

load('../../papers/short/ld5.Rdata')
plotIndv(out$samples,'Estimated Effects',ylim=c(-60,200))
lines(colMeans(out$samples), col = "orangered2", lwd = 1.2)
points(colMeans(out$samples), col = "orangered2", pch = 20)
```

## Results for "Does Everyone" Analysis

\vspace*{.2cm}

- Preferred model: **Analog representation** model

- Preferred `r printnum(out$BF[1])`-to-1 over the **None of the above** model

- Preferred `r "$3 \\times 10^{55}$"`-to-1 over the **Propositional representation** model

- Bayes factor for **Priming + spreading activation** model cannot be estimated 

\centering \includegraphics[width = 200px]{pics/skala.png}

```{r bf-5, fig.height=4.5, eval = F}
par(mar = c(3, .1, 2.5, 0))
layout(matrix(1:4
              , ncol = 2, byrow = T))

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Analog Representation"), side = 3, cex = 1.8)
mtext(bquote(BF["1u"] ~ "=" ~ .(round(out$BF[1], 2))), side = 3, cex = 1.7, line = -5)
box(col = "orangered2", lwd = 1.5)

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[2]: ~ "Priming + Spreading Activation"), side = 3, cex = 1.8)
mtext(bquote(BF["2u"] ~ "=" ~ 10^-55), side = 3, cex = 1.7, line = -5)

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[3]: ~ "Propositional Representation"), side = 3, cex = 1.8)
mtext(bquote(BF["3u"] %~~% 0), side = 3, cex = 1.7, line = -5)

plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[4]: ~ "None of the above"), side = 3, cex = 1.8)
mtext(bquote(BF["uu"] ~ "=" ~ 1), side = 3, cex = 1.7, line = -5)
```

## Take-away for the Pragmatic Psychologist

\vspace*{.5cm}

1. Theoretical ordinal predictions may account for everyone
2. "Everyone Does" models are parsimonious descriptions of common mechanisms
3. Bayesian analysis is the perfect tool for assessing multiple simultaneous ordinal predictions

\vspace*{1.5cm}

\begin{figure}
\hfill
\includegraphics[width = 80px]{pics/psychology.png}
\end{figure}

## Further Recources for the Pragmatic Psychologist

\vspace*{1cm}

- Accessible book to get started with systems of orders: @Hoijtink:2012

- Software packages: `bain` or `BayesFactor` for `R` users

- "Does everyone" analysis: @Haaf:Rouder:2017

- **Presentation and all code is available at:** \color{blue} \href{https://tinyurl.com/OrdinalConstraint}{tinyurl.com/OrdinalConstraint}

## Thank You

\vspace*{1cm}

\begin{figure}
\centering
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width = 110px]{pics/Fayette_Klaassen.jpg}
  
  Fayette Klaassen 
  
  (Utrecht University)
\end{minipage}%
\begin{minipage}{.5\textwidth}
  \centering
  \includegraphics[width = 170px]{pics/Jeff_Rouder.jpg}
  
  Jeff Rouder 
  
  (University of California - Irvine)
\end{minipage}
\end{figure}


#Additional Slides

## Preferred model

\vspace*{1cm}

\begin{columns}
\column{.1\textwidth}
\column{.8\textwidth}

```{r, fig.width=8, fig.height=4, fig.align='center'}
myCol <- c('lightblue', 'lavender', 'darkseagreen1', "mistyrose1")
layout(matrix(1:2, ncol = 2))

names <- c("LL", "LH", "HL", "HH")
o <- 1:4

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

posmat <- matrix(c(.5, .2, .8, .5, .93, .5, .5, .2), ncol = 2)

plotmat(M
        , pos =posmat
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)

M <- matrix(nrow = length(o), ncol = length(o), data=0)
M[1, 2] <- ''
M[1, 3] <- ''
M[2, 4] <- ''
M[3, 4] <- ''

posmat <- matrix(c(.5, .2, .8, .5, .6, .5, .5, .2), ncol = 2)

plotmat(M
        , pos = posmat
        , name = names[c(4, 2, 3, 1)]
        , curve = 0
        , box.type="round"
        , box.size=.06
        , box.prop=.8
        , box.col = myCol[c(4, 2, 3, 1)]
        , arr.length=0
        , box.cex = 1.2
        , relsize = 1
        , shadow.size = 0.007
        )
# plot(0, type='n', axes=FALSE, ann=FALSE)
mtext(expression(M[1]: ~ "Consistent +"), side = 3, cex = 1.5)
```

\column{.1\textwidth}
\end{columns}

## References

\scriptsize 
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}



---
nocite: |
  @Hoijtink:2012,
  @Rouder:etal:2012, 
  @Ryan:etal:2013
  @Rouder:etal:2005a
...

